|

### 2024-09-19 manuscript/paper.Rmd
Edit to Introduction, which is mostly done. Edits to methods and results which need joint writing.

Ecological release theory suggests that a population’s niche should change when important species interactions (e.g., resource competition) are relaxed or removed (reviewed in @Herrmannetal2021). Removal is posited to create ecological opportunities—i.e., aspects of the niche become newly accessible, and the focal population shifts and/or expands its resource use to the new opportunity (@ParentandCrespi2009; @Herrmannetal2021). Ecological release may then be followed by adaptive morphological evolution (@ParentandCrespi2009; @Herrmannetal2021). 
For example, a population undergoing ecological release can experience disruptive selection on males and females stemming from intraspecific competition over newly accessible resources (@BolnickandDoebeli2003; @BolnickandLau2008; @Cooperetal2011). This is predicted to result in intersexual divergence in habitat use and associated phenotypes (@Schoener1968; @Shine1989; @BolnickandDoebeli2003; @Butleretal2007; @BolnickandLau2008; @Cooperetal2011; but see @Stuartetal2021; @Blain2022). Such competition-driven “intraspecific character displacement” between the sexes is therefore one explanation for the evolution of sexual dimorphism (@PfennigandPfennig2012; @DeLisleandRowe2015; @DeLisleandRowe2017, @DeLisleetal2018).

Much of the theory and empirical data for character displacement between the sexes is based on release from resource competition specifically (e.g., @BolnickandDoebeli2003; @Cooperetal2011; @PfennigandPfennig2012; @DeLisleandRowe2015; @DeLisleetal2018). However, release from predation should also result in ecological release followed by the evolution of increased sexual dimorphism because the absence of predators should generate ecological opportunity (@ReimchenandNosil2004; @ParentandCrespi2009; @Herrmannetal2021).

Here, we tested the prediction that a release from predation should result in the evolution of increased sexual dimorphism using a well-preserved, finely-resolved sequence of a fossil threespine stickleback fish (Gasterosteus doryssus). The sequence in this depositional environment is comprised of two lineages (@Bell2009; @Cerasonietal2024). Lineage I was a low-armor form with zero to one dorsal spines and highly reduced pelvises, on average. It lasted for at least 93,000 years before it was replaced suddenly by a second lineage (@Belletal1985). Lineage II appeared in the depositional environment fully armored, with complete pelvic girdles, two pelvic spines, and three dorsal spines, on average (@Belletal2006; @Stuartetal2020). It then immediately began evolving adaptive reduction in its armor traits (@Huntetal2008; @Stuartetal2020) until it reached the same low-armor state previously held by lineage I. 
The observation of 93,000 years of low-armor stasis in lineage I (@Belletal1985) and the observation of rapid evolution of armor loss by lineage II, both suggest that this depositional environment lacked piscivorous fish (e.g., trout and other salmonids known to prey on modern threespine stickleback). Armor maintenance in extant threespine stickleback (Gasterosteus aculeatus) correlates strongly the presence of vertebrate piscivores; populations with less predation pressure typically have less armor (@Reimchen1994, @ReimchenandNosil2004; @Belletal1993; @Roestietal2023). Consistently, only three fossil trout have been recorded from this paleolake basin, sampled from the same rock that has revealed >15,000 threespine stickleback fossils as well as occasional killifish (Fundulus nevadensis) (@Bell2009; @Cerasonietal2024). Finally, tooth wear data from the lineage II sequence suggest that individuals in this lineage began eating more planktonic prey over time, expanding toward an open-water niche from the benthic, bottom-feeding niche they started with (Purnell et al. 2007). That this expansion into open water by lineage II coincided with armor reduction is further indicative of a limnetic system with fewer salmonid piscivores (@SchluterandMcPhail1992; @VamosiandSchluter2004; @Roestietal2023). In sum, we infer that lineage II likely started in a nearby paleolake basin that had predators since it was armored when it arrived (@Bell2009; @Cerasonietal2024). Lineage II then experienced release from predators and dietary expansion in the focal paleolake basin, generating a scenario consistent with the starting conditions for models predicting the evolution of increased sexual dimorphism following ecological release.

@@

For our study, we inferred sex in G. doryssus fossils by comparing multivariate morphological trait data from lineage II samples to the same multivariate trait set collected from multiple populations in the closely-related, extant Threespine Stickleback species complex (Gasterosteus aculeatus). Crucially, we determined G. aculeatus sex directly via dissection and/or PCR genotyping. This enabled us to use Multiple Imputation by Chained Equations (MICE; @MICE) to build a predictive multiple imputation algorithm (@little2002statistical) based on the multivariate morphology of G. aculeatus individuals of known sex. We applied this algorithm to the fossil data to impute individual sex based on morphology, treating fossil sex as a missing variable. With sex assigned to fossil specimens (and accounting for uncertainty in those assignments), we fit a modified Ornstein-Uhlenbeck (OU) (@OUProcess) model using a Bayesian framework to test for evolution of sexual dimorphism in each trait over ~16,000 years of lineage II. We hypothesized that sexual dimorphism should increase following ecological release from predators.

@@

## Fossil Specimen Data
We used Gasterosteus doryssus data that were previously reported by @Stuartetal2020, @Vojeetal2022, and @Siddiquietal2024. Briefly, the data were collected from fossil Series K from Quarry D (@Cerasonietal2024), dug from an open pit diatomite mine at 9.526° N, 119.094° W, near Fernley, Nevada, USA. Series K consisted of 18 samples taken at ~1000-year intervals, and mean sample times span ~16,363 years. Fish from series K were measured for 16 ecomorphological traits related to armor, swimming, and feeding (Table 1). Series K started at a previously documented horizon when a low armored lineage of stickleback with zero to one dorsal spines, zero pelvic spines, and highly reduced pelvises (i.e., lineage I described above) was completely replaced by a high armored lineage of stickleback with three dorsal spines, two pelvic spines, and complete pelvises (i.e., lineage II described above) (@Belletal2006; @Bell2009; @Stuartetal2020). Lineage II subsequently evolved reduction in armor, body size, and traits related to swimming and feeding (@Belletal2006; @Stuartetal2020; @Siddiquietal2024). The tempo and mode of armor reduction during this sequence suggests adaptive evolution by natural selection (@Huntetal2008), and we focus on the multivariate evolution of sexual dimorphism by this second lineage.

@@

## Extant Specimen data
To span the gamut of stickleback diversity for our predictive imputation model, we sampled modern stickleback from lakes containing generalist stickleback populations (@Hendryetal2009; @Bolnick2011) and from lakes containing benthic-limnetic species pairs (@Baumgartneretal1988; @SchluterandMcPhail1992). The generalist populations used here were the remainder of collections made by YES in 2013 and were previously described in @Stuartetal2017 (Table S1). These samples were fixed in formalin, then stained for bone with Alizarin Red in 2013. Benthic and limnetic specimens were kindly loaned by D. Schluter and his lab at University of British Columbia. They collected benthic and limnetic individuals from Enos Lake in 1988 and from Emily Lake, Little Quarry Lake, Paxton Lake, and Priest Lake in 2018 (Table S1). The Enos specimens had been fixed whole in formalin and stored in 40% isopropanol. The specimens from the other lakes were initially preserved whole in 95% ethanol in the field before being gradually transferred to water then formalin in the lab and ultimately stored in 40% isopropanol. In 2019, we stained these specimens for bone using Alizarin Red.
We next replicated fossil data collection (Table 1) on these extant specimens. Standard length as well as pelvic-spine length on each side were measured with calipers. We used a dissection microscope to count dorsal spines, pelvic spines, dorsal-fin rays, and anal-fin rays. Right and left-side pelvic girdle lengths and ectocoracoid lengths were measured from ventral photographs taken using a Canon EOS Rebel T7 with a Tamron 16-300 mm MACRO lens mounted on a leveled Kaiser RS1 copy stand. Specimens were held in place for ventral photographs using a small tabletop vise with an attached scale bar. Lateral X-rays were used to measure dorsal spine length, number of pterygiophores anterior to the pterygiophore holding the third spine, length of the pterygiophore just anterior to the third spine, cleithrum length, and pre-maxilla ascending branch length. We also counted vertebrae from the X-rays: abdominal vertebrae were counted anterior to the first vertebra with a haemal spine contacting an anal fin pterygiophore. Caudal vertebrae were posterior, including the first vertebra with the haemal spine contacting the anal fin pterygiophore (following @Aguirreetal2014). X-rays were taken with an AXR Hot Shot X-ray Machine (Associated X-ray Corporation) at the Field Museum of Natural History. Specimens were exposed at 35kV and 4mA. Small fish were exposed for 7s, medium fish for 8s, and large fish for 10s. We developed the film and scanned individual images of each fish using the B&W Negatives setting on an Epson Perfection 4990 Photo flatbed at 2400 dpi. Measurements from photographs and X-rays were taken with FIJI (@Schindelinetal2012) and its plugin ObjectJ (https://sils.fnwi.uva.nl/bcb/objectj/).

@@

To check for outliers, we calculated within-group means and standard deviations for each trait separately for K series fossil specimens (pooled across samples) and for extant specimens (within generalist, benthic, or limnetic categories). We noted trait values greater than 3.0 standard deviations from the mean as potential outliers. We deemed 3.0 s.d. to be a reasonable threshold for detecting errors without excluding biologically relevant values. We checked whether these potential outliers were a result of data entry and collection error and corrected them if they were. We turned the remaining outlier trait values to NAs. We size-corrected continuous traits only, as they varied with size, unlike count traits that are fixed during early development. We regressed each trait on standard length using a mixed-model regression, pooling all specimens, following @Stuartetal2017. We appended the size corrected data to the uncorrected trait data frame and used all of our data to build the imputation model described next.

## Missing data imputation, including fossil sex
We used multiple imputation (@little2002statistical) to impute the sex of the fossils many times.  Imputations were performed using the MICE (@MICE) algorithm and implemented in R (@R2022language) with M = 100 imputations performed. After imputing the sex of the fossil data, each completed data set is then used to fit a modified Ornstein–Uhlenbeck (OU) (@OUProcess) model using a Bayesian framework to look for sexual dimorphism across a variety of stickleback phenotypes.

@@

We impute the missing sex for the fossil data by sampling from the posterior predictive distribution $P(\boldsymbol{W}_{fossil}|\boldsymbol{W}_{extant},\boldsymbol{Y})$ using the multiple imputation by chained equations (MICE) algorithm (@MICE) with predictive mean matching.  Traditionally, the choice for the number of completed dats sets is a relatively small number such as $M = 5$ or $M = 10$.  However, @ZhouReiter2010 recommend a larger number of imputed data sets if the data users intend on performing Bayesians analysis after imputation, which in this case, we do.  Therefore, the imputation algorithm is run to obtain a total of $M = 100$ completed datasets. In addition to this, @ZhouReiter2010 suggests rather than using Rubin's combining rules to combine across imputed data sets, instead pool all of the draws from the posterior distributions across all of the imputed data sets to estimate the posterior distributions of parameters of interest.  We proceed with our Bayesian analysis in this manner.

\textcolor{red}{Akhil's stuff about validating the imputation model goes here.}

@@

## Completed Data Model 
\textcolor{red}{We should note somewhere that we are only using the $W_{fossil}$ in the modeling part.  We drop the $W_{extant}$.  So just note that $W_{ti}$ is really $W_{fossil,ti}$.  Not sure how to say this, but we need to make it clear that we are only using the fossil data for the OU modeling.  Greg, check and see if this makes sense.}

\textcolor{red}{I think we need to move a lot of these details to the appendix.  }

For a given imputed dataset, let $W_{ij}$ be the imputed sex and $\boldsymbol{Y}_{ij}$ be the $K \times 1$ vector of phenotypes for stickleback fossil $j$ at time $t_i$ where $i = 1, \ldots, T$ and $j = 1,\ldots,n_{t}$. Note, we are inputting only the fossil data into our model; connecting to the previous section, $W_{ij}$ and $\boldsymbol{Y}_{ij}$ can be interpretted as $W_{fossil,ij}$ and $\boldsymbol{Y}_{fossil,ij}$. In addition, we denote $Y_{K,ij}$, the last variable in $\boldsymbol{Y}_{ij}$, to be the standard length of the fish,
\begin{align}
{Y}_{K,ij} & \iid \left\{\begin{array}{lll} \mathcal{N}(\mu_{K,ft_i},&\sigma_{K}^2), & W_{ij} = \text{Female} \\ \mathcal{N}(\mu_{K,mt_i},&\sigma_{K}^2), & W_{ij} = \text{Male} \end{array}\right..
\label{eq:stl}
\end{align}

@@

Because we are fitting a dataset with a stochastic structure on the means of the phenotypes, and we are interested in the distribution of the overall structure of these means, we analyze the data via a Bayesian analysis.  Bayesian data analysis is also more naturally used when we have to impute data (\textbf{CITATION}). To aid in the sampling procedure for the discrete phenotypes, we declare

@@

# Results {#sec:results}
As previously described, four models for the mean process were considered for each trait.  Within each trait, the model with the lowest DIC was chosen for that train.  All subsequent results presented here use the model with the lowest DIC individually by trait.  Table \ref{whichmodel} shows the model that was selected for each trait.  About half of the traits had models with a trend, and most models did not include an OU component.  The traits that did include an OU component were caudal vertebra (mcv), pelvic spine lentgh (lpt), abdominal vertebra (mav), and standard length (stl) with caudal vertebra (mcv) being the only model with both a trend and an OU component.  This implies that not only does caudal vertebra size have a linear trend across sex, but the average size at one time period is highly correlated with the average size at the next time period.<!-- NOTE FOR MATT: You should mention something about MCV being the only one with a trend and an OU.  Explain why this might be and what it means.  -->

\textcolor{red} {
Yoel's interpretation of the above paragraph. Taking the model with the lowest DIC score as the best model for each trait, we found that cle, ds1, lps, and pmx did not fit either the trend model or the OU model for male-female trait mean differences. In contrast, we found a trend through time in male-female trait mean differences (i.e., sexual dimorphism) for mcv, ds2, ds3, ect, maf, and mdf. This suggests that these traits changed their dimorphism from one time period to the next, with an overall tendency toward ???GREATER/LESS??? dimorphism.
}
\textcolor{red} {
We found evidence for an OU process but no trend for lpt, mav, and stl, suggesting that these traits varied through time, about a fixed value for dimorphism. Finally, we found evidence that change in mcv fitted both a trend an an OU process, meaning ??? WHAT BIOLOGICALLY ???.
}

@@

# Discussion, Future work and conclusions {#sec:conclusions}
We predicted that release from predators would result in niche expansion and increased sexual dimorphism, based on several studies of modern stickleback. For example, in lakes where sculpin competitors are absent and stickleback (Roesti et al. 2023)
See Spoljaric and Reimchen 2008, page 512 right column for references and discussion of differences between benthic males and limnetic females. Male stickleback are benthic and littoral (Wootton 1976)…. Reimchen papers in general good for this section.

>>>>

Ecological release theory suggests that a population’s niche should change when important species interactions like resource competition or predation are relaxed or removed (reviewed in @Herrmannetal2021). Removal is posited to create ecological opportunity—i.e., aspects of the niche become newly accessible, and the focal population shifts and/or expands its resource use new resources (@ParentandCrespi2009; @Herrmannetal2021). Ecological release may then be followed by adaptive morphological evolution as traits change to reflect the new niche (@ParentandCrespi2009; @Herrmannetal2021).

For example, a population undergoing ecological release can experience disruptive selection on males and females stemming from intraspecific competition over newly accessible resources (@BolnickandDoebeli2003; @BolnickandLau2008; @Cooperetal2011). This is predicted to result in intersexual divergence in habitat use and associated phenotypes (@Schoener1968; @Shine1989; @BolnickandDoebeli2003; @Butleretal2007; @BolnickandLau2008; @Cooperetal2011; but see @Stuartetal2021; @Blain2022). Such competition-driven “intraspecific character displacement” between the sexes is therefore one explanation for the evolution of sexual dimorphism (@PfennigandPfennig2012; @DeLisleandRowe2015; @DeLisleandRowe2017, @DeLisleetal2018).

Much of the theory and empirical data for character displacement between the sexes is based on release from resource competition specifically (e.g., @BolnickandDoebeli2003; @Cooperetal2011; @PfennigandPfennig2012; @DeLisleandRowe2015; @DeLisleetal2018). However, release from predation might also drive the evolution of increased sexual dimorphism because an absence of predators should generate ecological opportunity (@ReimchenandNosil2004; @ParentandCrespi2009; @Herrmannetal2021).

Here, we tested the prediction that a release from predation results in the evolution of increased sexual dimorphism using a well-preserved, finely-resolved sequence of a fossil threespine stickleback fish (Gasterosteus doryssus). The sequence in this depositional environment is comprised of two lineages (@Bell2009; @Cerasonietal2024). Lineage I was a low-armor form with zero to one dorsal spines and highly reduced pelvises, on average. Lineage I lasted for at least 93,000 years before it was replaced by a second lineage, suddenly, likely on the order of years (@Belletal1985; Stuart et al. unpublished data). Lineage II appeared in the depositional environment fully armored, with complete pelvic girdles, two pelvic spines, and three dorsal spines (@Belletal2006; @Stuartetal2020). It then immediately began evolving adaptive reduction in its armor traits (@Huntetal2008; @Stuartetal2020) until it reached the same low-armor state previously held by lineage I.

The observation of 93,000 years of low-armor stasis in lineage I (@Belletal1985) and the observation of rapid evolution of armor loss by lineage II, both suggest that this depositional environment lacked piscivorous fish like trout and other salmonids known to prey on modern threespine stickleback. Armor presence in extant threespine stickleback (Gasterosteus aculeatus) correlates strongly with the presence of vertebrate piscivores; populations with less predation pressure typically have less armor (@Reimchen1994, @ReimchenandNosil2004; @Belletal1993; @Roestietal2023). In our paleolake basin, only three fossil trout have been found in the same section of rock that has revealed >20,000 threespine stickleback fossils as well as occasional killifish (Fundulus nevadensis) (@Bell2009; @Cerasonietal2024).  Thus, we reconstruct an evolutionar history in which it is likely that lineage II migrated from a nearby paleolake basin that had predators since it was armored when it arrived (@Bell2009; @Cerasonietal2024). Lineage II then experienced release from predators in the focal paleolake basin, generating the initial conditions of evolutionary models predicting the evolution of increased sexual dimorphism following ecological release. We test this prediction here.

@@

For our study, we inferred sex in G. doryssus fossils by comparing multivariate morphological trait data from lineage II samples to the same multivariate trait set collected from multiple populations in the closely-related, extant Threespine Stickleback species complex (Gasterosteus aculeatus). Crucially, we determined G. aculeatus sex directly via dissection and/or PCR genotyping. This enabled us to use Multiple Imputation by Chained Equations (MICE; @MICE) to build a predictive multiple imputation algorithm (@little2002statistical) based on the multivariate morphology of G. aculeatus individuals of known sex. We applied this algorithm to the fossil data to impute individual sex based on morphology, treating fossil sex as a missing variable. With sex assigned to fossil specimens, we fit a modified Ornstein-Uhlenbeck (OU) (@OUProcess) model using a Bayesian framework that accounts for uncertainty in sex classification to test for evolution of sexual dimorphism in each trait over ~16,000 years of lineage II.

@@

## Fossil Specimen Data
We used Gasterosteus doryssus data that were previously reported by @Stuartetal2020, @Vojeetal2022, and @Siddiquietal2024. Briefly, the data were collected from fossil Series K from Quarry D (@Cerasonietal2024), dug from an open pit diatomite mine at 9.526° N, 119.094° W, near Hazen, Nevada, USA. Series K consisted of 18 samples taken at ~1000-year intervals, and mean sample times span ~16,363 years. Fish from series K were measured for 16 ecomorphological traits related to armor, swimming, and feeding (Table 1). Series K started at the previously documented horizon when lineage I was replaced by lineage II. The tempo and mode of lineage II armor reduction during this sequence suggests adaptive evolution by natural selection (@Huntetal2008), and we focus on the multivariate evolution of sexual dimorphism by this second lineage.

@@

## Extant Specimen data
To span the gamut of stickleback diversity for our predictive imputation model, we sampled modern stickleback from lakes containing generalist stickleback populations (@Hendryetal2009; @Bolnick2011) and from lakes containing benthic-limnetic species pairs (@Baumgartneretal1988; @SchluterandMcPhail1992). The generalist specimens used here were collected by YES in 2013 and were previously described in @Stuartetal2017 (Table S1). These samples were fixed in formalin, then stained for bone with Alizarin Red in 2013. Benthic and limnetic specimens were kindly loaned by D. Schluter and S. Blain at the University of British Columbia. The Schluter lab collected benthic and limnetic individuals from Enos Lake in 1988 and from Emily Lake, Little Quarry Lake, Paxton Lake, and Priest Lake in 2018 (Table S1). The Enos specimens had been fixed whole in formalin and stored in 40% isopropanol. The specimens from the other lakes were initially preserved whole in 95% ethanol in the field before being gradually transferred to water then formalin in the lab and ultimately stored in 40% isopropanol. In 2019, we stained these specimens for bone using Alizarin Red.
We next replicated fossil data collection (Table 1) on these extant specimens. Standard length as well as pelvic-spine length on each side were measured with calipers. We used a dissection microscope to count dorsal spines, pelvic spines, dorsal-fin rays, and anal-fin rays. Right and left-side pelvic girdle lengths and ectocoracoid lengths were measured from ventral photographs taken using a Canon EOS Rebel T7 with a Tamron 16-300 mm MACRO lens mounted on a leveled Kaiser RS1 copy stand. Specimens were held in place for ventral photographs using a small tabletop vise with an attached scale bar. Lateral X-rays were used to measure dorsal spine length, number of pterygiophores anterior to the pterygiophore holding the third spine, length of the pterygiophore just anterior to the third spine, cleithrum length, and pre-maxilla ascending branch length. We also counted vertebrae from the X-rays: abdominal vertebrae were counted anterior to the first vertebra with a haemal spine contacting an anal fin pterygiophore. Caudal vertebrae were posterior, including the first vertebra with the haemal spine contacting the anal fin pterygiophore (following @Aguirreetal2014). X-rays were taken with an AXR Hot Shot X-ray Machine (Associated X-ray Corporation) at the Field Museum of Natural History. Specimens were exposed at 35kV and 4mA. Small fish were exposed for 7s, medium fish for 8s, and large fish for 10s. We developed the film and scanned individual images of each fish using the B&W Negatives setting on an Epson Perfection 4990 Photo flatbed at 2400 dpi. Measurements from photographs and X-rays were taken with FIJI (@Schindelinetal2012) and its plugin ObjectJ (https://sils.fnwi.uva.nl/bcb/objectj/).

@@

To check for outliers, we calculated within-group means and standard deviations for each trait separately for K series fossil specimens (pooled across samples) and for extant specimens (within generalist, benthic, or limnetic categories). We noted trait values greater than 3.0 standard deviations from the mean as potential outliers. We deemed 3.0 s.d. to be a reasonable threshold for detecting errors without excluding biologically relevant values. We checked whether these potential outliers were a result of data entry and collection error and corrected them if they were. We turned the remaining outlier trait values to NAs. We size-corrected continuous traits only, as they varied with size, unlike count traits that are fixed during early development. We regressed each trait on standard length using a mixed-model regression, pooling all specimens, following @Stuartetal2017. We appended the size corrected data to the uncorrected trait data frame and used all of our data (raw and size-corrected) to build the imputation model described next.

## Missing data imputation, including fossil sex
Briefly, We used multiple imputation (@little2002statistical) to impute the sex of the fossils.  Imputations were performed using the MICE (@MICE) algorithm and implemented in R (@R2022language) with M = 100 imputations. After imputing the sex, we fit a modified Orenstein-Uhlenbeck (OU; @OUProcess) model to each imputed data set using a Bayesian framework to {THIS NEEDS MORE}. We then quantified sexual dimorphism as the {THIS NEEDS MORE; MAYBE WE CAN WRITE TOGETHER}. We describe the details of the imputation model and OU modeling next.

@@

\textcolor{red}{AKHIL'S VALIDATION OF THE IMPUTATION MODEL GOES HERE}

We impute missing sex for the fossil data by sampling from the posterior predictive distribution $P(\boldsymbol{W}_{fossil}|\boldsymbol{W}_{extant},\boldsymbol{Y})$ using the multiple imputation by chained equations (MICE) algorithm (@MICE) with predictive mean matching.  Traditionally, the choice for the number of completed dats sets is a relatively small number such as $M = 5$ or $M = 10$.  However, @ZhouReiter2010 recommend a larger number of imputed data sets if users intend on performing Bayesian analysis after imputation.  Therefore, we imputed $M = 100$ complet datasets.

\textcolor{red}{presumably this also imputes missing data in the other traits? We should probably say that explicitly. -YS}

We then pooled all of the draws from the posterior distribution across all the imputed data sets to estimate the posterior distributions of our parameters of interest (i.e., the trait values), rather than usingRubin's combining rules, following @ZhouReiter2010. We used these posterior distributions in our Bayesian analysis.

@@

## Bayesian OU modeling
For Bayesian modeling, we dropped the extant data and focused on modeling the evolution of sexual dimorphism through time in the fossils.

\textcolor{red}{DOES THE NEXT SENTENCE CONTRADICT THE PREVIOUS ONE? BELOW, WE SAY THAT WE USED EACH IMPUTED DATA SET. ABOVE, WE SAY THAT WE POOLED ACROSS IMPUTED DATA SETS.-YS}
For a given imputed dataset, let $W_{ij}$ be the imputed sex and $\boldsymbol{Y}_{ij}$ be the $K \times 1$ vector of phenotypes for stickleback fossil $j$ at time $t_i$ where $i = 1, \ldots, T$ and $j = 1,\ldots,n_{t}$. Note, we are inputting only the fossil data into our model; connecting to the previous section, $W_{ij}$ and $\boldsymbol{Y}_{ij}$ can be interpreted as $W_{fossil,ij}$ and $\boldsymbol{Y}_{fossil,ij}$. In addition, we denote $Y_{K,ij}$, the last variable in $\boldsymbol{Y}_{ij}$, to be the standard length of the fish,
\begin{align}
{Y}_{K,ij} & \iid \left\{\begin{array}{lll} \mathcal{N}(\mu_{K,ft_i},&\sigma_{K}^2), & W_{ij} = \text{Female} \\ \mathcal{N}(\mu_{K,mt_i},&\sigma_{K}^2), & W_{ij} = \text{Male} \end{array}\right..
\label{eq:stl}
\end{align}

@@

Because we are fitting a dataset with a stochastic structure on phenotype means, and we are interested in the distribution of the overall structure of these means, we analyze the data via a Bayesian analysis <THIS LAST CLAUSE DOESN'T FOLLOW FROM THE BECAUSE CLAUSE. BETTER EXPLAIN WHY BAYES IS GOOD HERE>.  Bayesian data analysis is also more naturally used when we have to impute data (\textbf{CITATION}). To aid in the sampling procedure for the discrete phenotypes, we declare

@@

# Results {#sec:results}
<WE NEED RESULTS SECTION DESCRIBING THE RESULTS OF AKHIL'S MULTIPLE IMPUTATION, BOTH THE TESTS OF HOW WELL THE MODEL WORKED CLASSIFYING MODEL SAMPLES, AND SUMMARY STATISTICS OF FOSSIL IMPUTATION RESULTS -YS>

Within each trait, the model with the lowest DIC was chosen for that trait <THIS IS METHODS--WHERE IS THIS DESCRIBED IN THE METHODS>.  All subsequent results presented here use the model with the lowest DIC individually by trait <THIS IS METHODS--WHERE IS THIS DESCRIBED IN THE METHODS>.

Table \ref{whichmodel} reports the results of model fitting.  X of the Y traits were best fit by a trend model. Best models included an OU component for caudal and abdominal vertebra number (mcv, mav), pelvic spine length (lpt <LPT IS PTERYGIOPHORE LENGTH>) and standard length (stl). Caudal vertebra number was the only trait best fit by a model including both a trend and an OU component.  This implies that not only does caudal vertebra size have a linear trend across sex <WHAT DO YOU MEAN BY LINEAR TREND ACROSS SEX?>, but the average size at one time period is highly correlated with the average size at the next time period <I'VE ALWAYS FOUND THAT THIS AID TO MODEL INTEPRETATION IS BETTER IN THE METHODS, JUST AFTER THE MODELS ARE DESCRIBED>.

\textcolor{red} {
Yoel's interpretation of the above paragraph. Taking the model with the lowest DIC score as the best model for each trait, we found that cle, ds1, lps, and pmx did not fit either the trend model or the OU model for male-female trait mean differences. In contrast, we found a trend through time in male-female trait mean differences (i.e., sexual dimorphism) for mcv, ds2, ds3, ect, maf, and mdf. This suggests that these traits changed their dimorphism from one time period to the next, with an overall tendency toward ???GREATER/LESS??? dimorphism. IS THIS CORRECT? - YS
}
\textcolor{red} {
We found evidence for an OU process but no trend for lpt, mav, and stl, suggesting that these traits varied through time, about a fixed value for dimorphism. Finally, we found evidence that change in mcv fitted both a trend an an OU process, meaning ??? WHAT BIOLOGICALLY ???.
}

@@

# Discussion, Future work and conclusions {#sec:conclusions}
We predicted that release from predators would result in niche expansion and increased sexual dimorphism, based on several studies of modern stickleback. For example, in lakes where sculpin competitors are absent and stickleback (Roesti et al. 2023)
See Spoljaric and Reimchen 2008, page 512 right column for references and discussion of differences between benthic males and limnetic females. Male stickleback are benthic and littoral (Wootton 1976)…. Reimchen papers in general good for this section.
Moreover, tooth wear data from the lineage II sequence suggest that individuals in this lineage began eating more planktonic prey over time, expanding toward an open-water niche from the benthic, bottom-feeding niche they started with (Purnell et al. 2007). That this expansion into open water by lineage II coincided with armor reduction is further indicative of a limnetic system with fewer salmonid piscivores (@SchluterandMcPhail1992; @VamosiandSchluter2004; @Roestietal2023).
